Unsupervised Learning for Identifying Events in Active Target Experiments

### 2020-08-06 doc/Manuscript/article.tex
Updates from Overleaf

Within the context of machine learning methods applied to the analysis of nuclear physics experiments, 
the purpose of this work is thus to explore the application of unsupervised learning algorithms to event identification from an active target detector. The necessity to identify events from raw data prior to full processing is becoming a major issue in the data analysis of detectors with complex responses such as the AT-TPC.  After these introductory words, we review the experimental information in the next section. In section \ref{sec:datadescript} we describe the set up of the data while sections \ref{sec:me

@@

To train the MIXAE clustering algorithm, we use the large simulated data set with $M=80000$ points, evenly distributed between proton- and carbon-events. The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events. On real data the algorithm is trained on an unlabelled set of data, and evaluated on an unlabelled subset\footnote{see table \ref{tab:data sets} for details}. Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search for the optimal configuration. Finally, for the best parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm.

>>>>

Within the context of machine learning methods applied to the analysis of nuclear physics experiments, 
the purpose of this work is thus to explore the application of unsupervised learning algorithms to event identification from an active target detector. The necessity to identify events from raw data prior to full processing is becoming a major issue in the data analysis of detectors with complex responses such as the AT-TPC.  After these introductory words, we review the experimental information in the next section. In section \ref{sec:datadescript} we describe the set up of the data while sections \ref{sec:methods} and \ref{sec:results}
present the methods applied and our results and discussions, respectively. Finally, our conclusions and perspectives for future work are presented in section \ref{sec:conclusion}.

@@

To train the MIXAE clustering algorithm, we use the large simulated data set with $M=80000$ points, evenly distributed between proton- and carbon-events. The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events. On real data the algorithm is trained on an unlabelled set of data, and evaluated on a labelled subset\footnote{see table \ref{tab:data sets} for details}. Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search for the optimal configuration. Finally, for the best parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm.
